DNA meth and IVF

### 2023-12-21 figs/GSE158422_pkgdev.Rmd
must update gene symbols

suppressPackageStartupMessages({
  library("limma")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("tictoc")
  library("mitch")
  library("kableExtra")
  library("beeswarm")
  library("missMethyl")
  library("gridExtra")
  library("png")
})

@@

Now try for efficient method to incorporate into the package.

```{r,anno2}

library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")

@@

>>>>

suppressPackageStartupMessages({
  library("limma")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("HGNChelper")
  library("tictoc")
  library("mitch")
  library("kableExtra")
  library("beeswarm")
  library("missMethyl")
  library("gridExtra")
  library("png")
})

@@

tic()
new.hgnc.table <- getCurrentHumanMap()
fix <- checkGeneSymbols(gt$gene,map=new.hgnc.table)
fix2 <- fix[which(fix$x != fix$Suggested.Symbol),]
length(unique(fix2$x))
gt$gene <- fix$Suggested.Symbol
toc()

@@

# fix gene symbols
tic()
fix <- checkGeneSymbols(gt2$gene,map=new.hgnc.table)
fix2 <- fix[which(fix$x != fix$Suggested.Symbol),]
length(unique(fix2$x))
gt2$gene <- fix$Suggested.Symbol
toc()

### 2023-12-21 figs/GSE158422_rna_int.Rmd
must update gene symbols

suppressPackageStartupMessages({
  library("DESeq2")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("tictoc")
  library("limma")
  library("mitch")
  library("kableExtra")
  library("beeswarm")
  library("missMethyl")
  library("HGNChelper")
  library("gplots")
  library("gridExtra")
  library("png")
})

@@

@@

@@

>>>>

suppressPackageStartupMessages({
  library("DESeq2")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("HGNChelper")
  library("tictoc")
  library("limma")
  library("mitch")
  library("kableExtra")
  library("beeswarm")
  library("missMethyl")
  library("HGNChelper")
  library("gplots")
  library("gridExtra")
  library("png")
})

@@

new.hgnc.table <- getCurrentHumanMap()
fix <- checkGeneSymbols(gt$gene,map=new.hgnc.table)
fix2 <- fix[which(fix$x != fix$Suggested.Symbol),]
length(unique(fix2$x))
gt$gene <- fix$Suggested.Symbol

@@

fix <- checkGeneSymbols(gb$gene,map=new.hgnc.table)
fix2 <- fix[which(fix$x != fix$Suggested.Symbol),]
length(unique(fix2$x))
gb$gene <- fix$Suggested.Symbol

@@

fix <- checkGeneSymbols(gp$gene,map=new.hgnc.table)
fix2 <- fix[which(fix$x != fix$Suggested.Symbol),]
length(unique(fix2$x))
gp$gene <- fix$Suggested.Symbol

### 2023-12-21 figs/GSE158422_sensitivity.Rmd
must update gene symbols

anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","UCSC_RefGene_Group","Islands_Name","Relation_to_Island")])

@@

tic()
anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","UCSC_RefGene_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt <- stack(gp2)
colnames(gt) <- c("gene","probe")
dim(gt)
str(gt)
toc() #8.735s

@@

## LMRM - limma mean rank mitch

tic()
anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","UCSC_RefGene_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt <- stack(gp2)
colnames(gt) <- c("gene","probe")
dim(gt)
str(gt)
toc() #8.735s

dm$rank <-  rank(dm$t) - nrow(subset(dm,t<0))
mm <- merge(dm,gt,by.x=0,by.y="probe")
head(mm)
mma <-aggregate(mm$rank ~ gene,mm,mean)
rownames(mma) <- mma$gene
mma$gene = NULL
colnames(mma) <- "meanrank"
lmrmres <- runmitch(m=mma,genesets=gs_symbols,cores=16)

lmrmsig <- subset(lmrmres,p.adjustANOVA<0.05)

top <- head(lmrmsig,20)
top <- top[order(top$s.dist),]

hist(lmrmsig$s.dist)

# distinguish up (TRUE) and down (FALSE)
lmrmsigsets <- paste(rownames(lmrmsig),as.character(lmrmsig$s.dist > 0))
lmrmsigsets2 <- gsub(" TRUE"," UP",lmrmsigsets)
lmrmsigsets2 <- gsub(" FALSE"," DN",lmrmsigsets2)

>>>>

anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","UCSC_RefGene_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt <- stack(gp2)
colnames(gt) <- c("gene","probe")
dim(gt)
str(gt)
head(gt)

new.hgnc.table <- getCurrentHumanMap()
fix <- checkGeneSymbols(gt$gene,map=new.hgnc.table)
fix2 <- fix[which(fix$x != fix$Suggested.Symbol),]
length(unique(fix2$x))
gt$gene <- fix$Suggested.Symbol

@@

@@

### 2023-12-21 figs/GSE158422_simulate020.Rmd
must update gene symbols

suppressPackageStartupMessages({
  library("stringi")
  library("limma")
  library("missMethyl")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library('org.Hs.eg.db')
  library("psych")
  library("mitch")
  library("kableExtra")
})

>>>>

suppressPackageStartupMessages({
  library("stringi")
  library("limma")
  library("missMethyl")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("HGNChelper")
  library('org.Hs.eg.db')
  library("psych")
  library("mitch")
  library("kableExtra")
})

### 2023-12-21 figs/aging_example.Rmd
must update gene symbols

suppressPackageStartupMessages({
  library("limma")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("tictoc")
  library("mitch")
  library("gplots")
  library("kableExtra")
  library("beeswarm")
  library("missMethyl")
  library("gridExtra")
  library("png")
})

@@

@@

>>>>

suppressPackageStartupMessages({
  library("limma")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("HGNChelper")
  library("tictoc")
  library("mitch")
  library("gplots")
  library("kableExtra")
  library("beeswarm")
  library("missMethyl")
  library("gridExtra")
  library("png")
})

@@

Update old gene symbols.

@@

tic()
new.hgnc.table <- getCurrentHumanMap()
fix <- checkGeneSymbols(gt$gene,map=new.hgnc.table)
fix2 <- fix[which(fix$x != fix$Suggested.Symbol),]
length(unique(fix2$x))
gt$gene <- fix$Suggested.Symbol
toc()

### 2023-12-21 figs/art_example.Rmd
must update gene symbols

@@

@@

>>>>

Update gene symbols.

@@

tic()
new.hgnc.table <- getCurrentHumanMap()
fix <- checkGeneSymbols(gt1$gene,map=new.hgnc.table)
fix2 <- fix[which(fix$x != fix$Suggested.Symbol),]
length(unique(fix2$x))
gt1$gene <- fix$Suggested.Symbol
toc()

@@

tic()
new.hgnc.table <- getCurrentHumanMap()
fix <- checkGeneSymbols(gt2$gene,map=new.hgnc.table)
fix2 <- fix[which(fix$x != fix$Suggested.Symbol),]
length(unique(fix2$x))
gt2$gene <- fix$Suggested.Symbol
toc()
